---
title: "Final_RMD"
author: "Paul Chong"
date: "12/16/2021"
output: html_document
---

```{r, include=FALSE}

# PACKAGES AND DEPENDENT CODE FROM CODE.RMD

library(readr)
library(knitr)
library(ggplot2)
library(dplyr)

library(ggpubr)


knitr::opts_chunk$set(dev = 'pdf')

```






# Calculated vs. Computed (cont.)

```{r}
loan <- 144000

drate <- 0.07 # default rate = 0.7
dloss <- -174115 # loss from each default
n_loans <- 100000

# Monte Carlo method 
B <- 1000 # Number of iterations 
results <- replicate(B,{ 
    
        temp <-sample(c(0,dloss),n_loans,
                      prob=c(1-drate, drate),
                      replace = TRUE) 
        sum(temp)})


```

# Calculated vs. Computed (cont.)
```{r, echo=FALSE}
cat("Calculation result:", n_loans * drate * dloss, "\n")
cat("Monte Carlo result:",mean(results))
```

# Calculated vs. Computed (cont.)

```{r,echo=FALSE, warning=FALSE}
hist(results,breaks=50,xlab="Profit in US dollars", 
   main="Histogram of Loss Pharmacist Student Loan Debt")
```

# Calculated vs. Computed (cont.)

+ So this begs the question, how much interest is needed to prevent a loss?

# Minimum Interest?

+ Our equation for profit: $Gain * (1 - Default_{Rate}) - (Default_{Loss} * Default_{Rate}) = Profit$

+ We want to test for profit of $0, to check for minimum interest that the bank has to charge

+ $Gain * (1 - Default_{Rate}) - (Default_{Loss} * Default_{Rate}) = 0$

+ $Gain = (Default_{Loss} * Default_{Rate}) / (1 - Default_{Rate})$

+ We can calculate interest as Interest = Gain/Loan

```{r,echo=FALSE, cache=TRUE}
#25% of the sudent debt, is help by active student

interest <- 0.000    # Start with a 0% interest

result <- 0

proportion <- c(1-drate,drate)
while (result <= 0){ # Each iteration increases the interest rate until the bank's profit is greater than $0.
    interest <- interest + 0.001 # Increase interest by 0.1%, to find the interest rate to to set the profit to zero 
    result <- mean(replicate(B,{ #Monte Carlo Method
                        temp <- sample(c(loan*interest,dloss), n_loans, prob = proportion, replace = TRUE)
                        
                        sum(temp)
    }))
}

cat("Interest from Simulation:", interest, "\n")

cat("Interest from Equation", -(dloss * drate / (1-drate))/loan)

```

# Minimum Interest? 

+ Again, the calculated and computed are very similar, let's see what the distribution of the bank's returns will look like with a 9.2% interest rate. 

```{r, echo = FALSE, cache=TRUE}
simu <- replicate(B,{temp <- sample(c(loan*0.091,dloss), n_loans, prob = proportion, replace = TRUE)
                    sum(temp)
    })


#mean(simu <=0)

```

# Minimum Interest?

```{r, echo = FALSE}
hist(simu,breaks=100,xlab="Profit in US dollars", 
   main="Histogram of Profit from Pharmacy Student")


```
We can see that the mean profit in this histogram is zero, representing $0 loss in 

Money loss chance of less than 5%

```{r , cacheTRUE}
set.seed(400)


#Since we we include the default rate is 9.1% or 0.091, so the z-score is this case will be -1.329 from the table (calculated from qnorm)
alpha <- qnorm(.091)
upper <- - dloss* (drate*n_loans-(alpha)*sqrt(n_loans*drate*(1-drate)))
lower <- (n_loans*(1-drate)+(alpha)*sqrt(n_loans*drate*(1-drate)))
gain <- upper/lower

# Interest formula
eq <- gain/loan

cat("Interest from Equation: ", eq)


```


```{r , cacheTRUE}
set.seed(400)

#when we have to include the percent taht student can't pay the moneyback

interest <- 0.0    # Start with a 0% interest

result <- 0

while (TRUE){
    interest <- interest + 0.001 # Each interation increases interest by 0.1% until the result is greater than $0.
    result <- replicate(B,{
                        temp <- sample(c(loan*interest,dloss), n_loans, prob = c(1-drate,drate), replace = T)
                        sum(temp)
    })
    
    if (mean(result <= 0)<=0.05){break} # Test if the scenarios where profit was < 0 represent the 5% of the total.
}

cat("The output interest from simulation is ", interest)




```
```{r, cacheTRUE}
set.seed(400)
lonVec <- c(loan*interest,dloss)
simu2 <- replicate(B,{temp <- sample(lonVec, n_loans, prob = proportion, replace = T)
                    sum(temp)
    })
hist(simu2,breaks=breaksto,xlab="Profit in US dollars", 
   main="Histogram of profit from Student Loans")



```

```{r }
set.seed(400)

h<-hist(Sample, breaks=breaksto, xlab="Profit in US dollars", 
   main="Histogram of profit from Student Loans") 
x<-seq(min(simu2),max(simu2),length=40) 
y<-dnorm(x,mean=mean(simu2),sd=sd(simu2)) 
y <- y*diff(h$mids[1:2])*length(simu2) 
lines(x, y, col="red", lwd=2)




```



```{r }

quantile_val <- pnorm(0.05,mean = mean(simu2), sd = sd(simu2))
cat("5th Quantile:",quantile_val )



```


```{r }

ccat = cut(h$breaks, c(-Inf, quantile_val, Inf))
plot(h, col=c("red","green")[ccat],xlab="Profit in US dollars", 
   main="Histogram of Profit from Scamming College Students")


```