---
title: "Final Paper 400"
author: "Paul Chong, Nat Koonmee, Mark Gardner"
date: "12/16/2021"
output: pdf_document
---

```{r, include=FALSE}

# PACKAGES AND DEPENDENT CODE FROM CODE.RMD


library(readr)
library(knitr)
library(ggplot2)
library(dplyr)

library(ggpubr)


knitr::opts_chunk$set(dev = 'pdf')

```

# Monte-Carlo Estimations of Pharmaceutical Student Loan Debt Default 


In a paper from the American Journal of Pharmaceutical Education, Mattingly and Ulbrich conducted a simulation on net income for pharmaceutical students. They report that the average amount of money borrowed by pharmaceutical students in 2014 is around 144,718 USD. That estimated amount has increased to around 179,514 USD, as reported by recent graduates in 2020. This finding was the main inspiration for our study. As university students ourselves, we wanted to understand how the industry uses Monte Carlo methods to determine the best estimates for making profit off student loans. In this paper, we will be analyzing the average amount that a bank would lose from defaults on loans. We will then use monte carlo simulations to determine the minimum interest rate to ensure that the bank won’t lose money. We will then use Monte Carlo to see if we can find an interest rate that would lower the chance of losing money to 5%.

## Methodology: 
We used monte carlo estimation to predict a bank’s exposure to pharmaceutical student loan default.  The total exposure to loss was calculated as the number of loans multiplied by the default rate multiplied by the default loss.

Defining our parameters, our simulated average loan was 144,000 USD, while our default rate and loss were 7% and 174,115 USD, respectively.  Our MC estimation consisted of 1000 iterations with the defined parameters, resulting in calculated (actual) and monte carlo estimations of total losses of  1,218,805,000 USD and 1,218,471,744 USD, respectively.  

Equation: $Loss = n_{loans} * Default_{Rate} * Default_{loss}$

This equation is the original one we used, and it doesn't attempt to get profit.  It's just calculating the amount the bank would lose if it charged 0% interest. 

```{r, include = FALSE, cache = TRUE}
loan <- 144000

drate <- 0.07 # default rate = 0.7
dloss <- -174115 # loss from each default
n_loans <- 100000

# Monte Carlo method 
B <- 1000 # Number of iterations 
results <- replicate(B,{ 
    
        temp <-sample(c(0,dloss),n_loans,
                      prob=c(1-drate, drate),
                      replace = TRUE) 
        sum(temp)})

cat("Calculation result:", n_loans * drate * dloss, "\n")
cat("Monte Carlo result:",mean(results))

```

The calculated results were found as described above, while the MC estimation of the student loan exposure was a result of sampling the given parameters iteratively, creating vectors of student default debt consisting of either 0 or the estimated loss: $174,115.  Each of these vectors of student default were 10,000 rows long, and our MC estimate consisted of taking 1000 of these 10,000 student samples. 


Our next step was to find the minimum interest rate required for the bank to expect 0 losses.  We used another MC estimation of 100 iterations to predict the minimum interest rate required for 0 expected losses, and we found the interest from simulation and from the actual equation to be 0.091 and 0.09100993, respectively.  Clearly, our monte-carlo estimates were good representations of the underlying theoretical population. 

Our equation for profit: $Gain * (1 - Default_{Rate}) - (Default_{Loss} * Default_{Rate}) = Profit$

We want to test for profit of $0, to check for minimum interest that the bank has to charge

$Gain * (1 - Default_{Rate}) - (Default_{Loss} * Default_{Rate}) = 0$

$Gain = (Default_{Loss} * Default_{Rate}) / (1 - Default_{Rate})$

This is the equation used by the bank to calculate the interst rate necessary to generate profit. 

We can calculate interest as Interest = Gain/Loan

```{r,echo=FALSE, cache=
