---
title: "Monte Carlo Methods for Scamming College Students"
author: "Paul Chong, Mark Gardner, Nat Koonmee"
date: "12/06/21"
output:
  slidy_presentation: default
  pdf_document: default
  ioslides_presentation: default
  beamer_presentation: default
---

```{r, include=FALSE}
# PACKAGES AND DEPENDENT CODE FROM CODE.RMD
library(readr)
library(knitr)
library(ggplot2)
library(dplyr)
library(ggpubr)
knitr::opts_chunk$set(dev = 'pdf')
```



# Motivation

As university students ourselves, a number of us are attending this wonderful class with borrowed money from the bank.

Our team wanted to play devil's advocate and find how banks can make profit from student loans

In this presentation, we will
+ Explain our inspiration behind this experiment
+ Provide background information on our experiment
+ Explain our replication
+ Further expand on our findings based on our statistical knowledge and 

# About the experiment inspiration

+ The inspiration for the experiment comes from a paper from the American Journal of Pharmaceutical Education.

+ "Evaluating the Financial Burdens for Graduating Pharmacists"

+ Compares income of medical practitioners between 2009 and 2014 using their income and student loan debt

# About the experiment inspiration (cont.)

+ We want to replicate part of the experiment using Monte Carlo and play devil's advocate for the bank

+ Our team wanted to see if the bank is able to make profit off of pharmaceutical students

# How much does the bank lose to defaults on student loans?

+ Average loan for pharmaceutical student is $144,000

+ Default rate is around 7%

+ Average loss of one person's default is $174,000

+ Number of loans handed out of medium-sized, private bank is a little under 100,000

# Calculated vs. Computed

+ We are going to calculate mathematically the amount of money the bank loses due to defaults, assuming the interest rate is 0%!

+ Bank is lending money for free

+ Equation: Loss = Number of Loans * Default Rate * Default Loss per Person

+ Loss = $1,218,805,000

# Calculated vs Computed (cont.)

+ In Monte Carlo method, we ran for 1000 iterations

+ Could have done more for better result, but didn't want to waste time on preliminaary methods. 

+ Monte Carlo Estimation Loss = $1,219,456,712

+ Both the calculated and computed estimates are very similar

# Calculated vs. Computed (cont.)

```{r, echo=FALSE}
loan <- 144000
default_rate <- 0.07
#approx around 7 years to pay all the loan back
default_loss <- -174115
n_loans <- 100000
# Monte Carlo method 
B <- 1000 # Number of iterations 
results <- replicate(B,{ 
    
        temp <-sample(c(0,default_loss),n_loans,prob=c(1-default_rate, default_rate),
                      replace = TRUE) 
        sum(temp)})
# Comparisson between calculated result and Monte Carlo result
cat("Calculation result:", n_loans * default_rate * default_loss, "\n")
cat("Monte Carlo result:",mean(results))
```

```{r,echo=FALSE}
hist(results,breaks=100,xlab="Profit in US dollars", 
   main="Histogram of Loss Pharmacist Student Loan Debt")
```

# Calculated vs. Computed (cont.)

+ So this begs the question, how much interest is needed to prevent a loss?

# Minimum Interest?

+ Our equation for profit: $Gain * (1 - Default_{Rate}) - (Default_{Loss} * Default_{Rate}) = Profit$

+ We want to test for profit of $0, to check for minimum interest that the bank has to charge

+ $Gain * (1 - Default_{Rate}) - (Default_{Loss} * Default_{Rate}) = 0$

+ $Gain = (Default_{Loss} * Default_{Rate}) / (1 - Default_{Rate})$

+ We can calculate interest as GInterest = ain/Loan

```{r,echo=FALSE, cache=TRUE}
#25% of the sudent debt, is help by active student
interest_rate <- 0.000    # Start with a 0% interest
result <- 0
proportion <- c(1-default_rate,default_rate)
while (result <= 0){ # Each iteration increases the interest rate until the bank's profit is greater than $0.
    interest_rate <- interest_rate + 0.001 # Increase interest by 0.1%, to find the interest rate to to set the profit to zero 
    result <- mean(replicate(B,{ #Monte Carlo Method
                        temp <- sample(c(loan*interest_rate,default_loss), n_loans, prob = proportion, replace = TRUE)
                        
                        sum(temp)
    }))
}
cat("Interest from Simulation:", interest_rate, "\n")
cat("Interest from Equation", -(default_loss * default_rate / (1-default_rate))/loan)
```

# Minimum Interest? 

+ Again, the calculated and computed are very similar, let's see what the distribution of the bank's returns will look like with a 9.2% interest rate. 

```{r, echo = FALSE}
simu <- replicate(B,{temp <- sample(c(loan*0.091,default_loss), n_loans, prob = proportion, replace = TRUE)
                    sum(temp)
    })
mean(simu <=0)
hist(simu,breaks=100,xlab="Profit in US dollars", 
   main="Profit from Pharmaceutical student loan default")
```


```{r, echo = FALSE}
total <- c()
  for(i in 1:30){# bc normally distributed by CLT? 
  simu <- replicate(B,{temp <- sample(c(loan*0.091,default_loss), n_loans, prob = proportion, replace = TRUE)
                      sum(temp)
      })

total[i] <- mean(simu <=0)

  }
mean9.1 <- mean(total)
mean9.1
hist(simu,breaks=100,xlab="Profit in US dollars", 
   main="Profit from Pharmaceutical student loan default")
```

```{r, echo = FALSE}
total <- c()
  for(i in 1:30){# bc normally distributed by CLT? 
  simu <- replicate(B,{temp <- sample(c(loan*0.09100993,default_loss), n_loans, prob = proportion, replace = TRUE)
                      sum(temp)
      })

total[i] <- mean(simu <=0)

  }
mean9.100993 <- mean(total)
mean9.100993
mean(simu <=0)
hist(simu,breaks=100,xlab="Profit in US dollars", 
   main="Profit from Pharmaceutical student loan default")
```


```{r, echo = FALSE}
total <- c()
  for(i in 1:30){# bc normally distributed by CLT? 
  simu <- replicate(B,{temp <- sample(c(loan*0.092,default_loss), n_loans, prob = proportion, replace = TRUE)
                      sum(temp)
      })

total[i] <- mean(simu <=0)

  }
mean9.2 <- mean(total)
mean9.2
mean(simu <=0)
hist(simu,breaks=100,xlab="Profit in US dollars", 
   main="Profit from Pharmaceutical student loan default")
```

# Minimum Interest?

+ There are problems with this graph, as theoretically the $0 in profit should be in the middle of the distribution

+ Still around 30%-40% chance bank loses money

+ However, there is still room for improvement

# Can we bring that frequency down to 5%?

+ What is the interest rate that can offer money loss frequency of less than 5%?

+ We hope to add this to the paper next week

+ We are assuming normal distribution










